---
title: 'Lab 3: Plotting Frequencies and Distributions'
format:
  html:
    self-contained: true
editor: source
knitr:
  opts_chunk:
    message: false
---

```{r}
#| document: show
library(spotifyr)
# id <- 'FILL CLIENT ID HERE'
# secret <- 'FILL CLIENT SECRET HERE'
# Sys.setenv(SPOTIFY_CLIENT_ID = id)
# Sys.setenv(SPOTIFY_CLIENT_SECRET = secret)
```

```{r}
#| warning: false
#| message: false
#| document: show
library(tidyverse)
spotify_playlists <- get_playlist_audio_features(
  username = "poiril",
  playlist_uris = c("7ryj1GwAWYUY36VQd4uXoq",
                    "3DYUw0nHB9o8tLZKQup4zp",
                    "37i9dQZF1DXdOEFt9ZX0dh"),
  authorization = get_spotify_access_token()
) %>%
  select(-c(track.artists, 
            track.available_markets, 
            track.album.artists, 
            track.album.available_markets, 
            track.album.images))
```

###### Question

::: question
How many songs are in each playlist? Create a plot to visualize this, and order the results by the number of songs. Be sure to give it a descriptive title and labels covering all 5 essential components of data context.
:::


```{r}
# Create plot here
```

###### Question

::: question
What is the distribution of valence across all of the songs (in intervals of 0.1 valence)? Create a histogram to visualize this. Be sure to give it a descriptive title and labels covering all 5 essential components of data context.
:::


```{r}
# Create plot here
```

###### Question

::: question
What is the distribution of valence across all of the songs (in intervals of 0.1 valence) *in each playlist*? Create a histogram to visualize this. Be sure to give it a descriptive title and labels covering all 5 essential components of data context.
:::


```{r}
# Create plot here
```

###### Question

::: question
What are differences in the summary statistics (max, min, median, etc.) of the valence of songs in each playlist? Create grouped boxplot to visualize this. Be sure to give it a descriptive title and labels covering all 5 essential components of data context.
:::


```{r}
# Create plot here
```

###### Question

::: question
Do happier songs tend to be more danceable *in each playlist*? Create a plot to visualize this. Be sure to give it a descriptive title and labels covering all 5 essential components of data context. Also be sure to adjust your plot to address overplotting.
:::


```{r}
# Create plot here
```

###### Question

::: question
Do songs composed with the minor or major `mode_name` tend to be happier *in each playlist*? Create a plot to visualize this. Be sure to give it a descriptive title and labels covering all 5 essential components of data context.
:::


```{r}
# Create plot here
```

###### Question

::: question
Do happier songs tend to have a higher tempo across all playlists? What role might the song's mode play? Create a plot to visualize this. Be sure to give it a descriptive title and labels covering all 5 essential components of data context. Also be sure to adjust your plot to address overplotting.
:::


```{r}
#Create plot here
```

###### Question

::: question
Based on all of your plots, summarize what you learned about the joyfulness of these three playlists below.
:::


