---
title: 'Lab 5: Exploratory Data Analysis'
format:
  html:
    self-contained: true
editor: source
knitr:
  opts_chunk:
    message: false
---

```{r}
#| warning: false
#| messages: false
#| document: show
library(tidyverse)

sqf_url <- "https://www1.nyc.gov/assets/nypd/downloads/zip/analysis_and_planning/stop-question-frisk/sqf-2011-csv.zip"
temp <- tempfile()
download.file(sqf_url, temp)
sqf_zip <- unzip(temp, "2011.csv")
sqf_2011 <- read.csv(sqf_zip, stringsAsFactors = FALSE) 
sqf_2011_race_cat <- read.csv("https://raw.githubusercontent.com/lindsaypoirier/STS-101/master/Data/SQF/sqf_race_categories.csv", stringsAsFactors = FALSE) 
rm(sqf_url)
rm(temp)
rm(sqf_zip)
```

```{r}
#| warning: false
#| messages: false
#| document: show
sqf_2011 <- 
  sqf_2011 |> 
  select(pct, race, age, frisked, pistol, riflshot, asltweap, knifcuti, machgun, othrweap, sumissue, arstmade) |>
  left_join(sqf_2011_race_cat, by = "race") |>
  mutate(across(frisked:arstmade, 
         ~ case_when(. == "Y" ~ 1, . == "N" ~ 0)))
rm(sqf_2011_race_cat)
```

###### Question

::: question

Add two new columns. The first should indicate whether a weapon was found, and the second should indicate whether an arrest/summons was made.

:::


```{r}
#| eval: false
sqf_2011 <- 
  sqf_2011 |>
  #Add a variable for weapon found
  _____(wpnfound = case_when(pistol == 1 |
                               riflshot == 1 | 
                               asltweap == 1 |
                               knifcuti == 1 | 
                               machgun == 1 | 
                               othrweap == 1 ~ 1,
                             TRUE ~ 0))
sqf_2011 <- 
  sqf_2011 |>
  #Add a variable for arrest made or summons issued
  _____(arrestsumm = case_when(sumissue == 1 | 
                                arstmade == 1 ~ 1,
                               TRUE ~ 0))
```

###### Question

::: question

Subset the dataset to the six variables listed in the data dictionary above.

:::


```{r}
#| eval: false
sqf_2011 <-
  sqf_2011 |>
  _____(pct, arrestsumm, _____, wpnfound, race_cat, _____)
```

###### Question

::: question
Calculate the number of stops in 2011. If you are not sure which function to use below, you may want to refer to the list of Summary functions in the the Data Wrangling cheatsheet. Remember that each row in the data frame is a stop.
:::


```{r}
#| eval: false
total_stops <-
  sqf_2011 |>
  summarize(Count = _____) |>
  pull()

total_stops
```

###### Question

::: question
How many stops did not result in an arrest or summons in 2011? What percentage of stops did not result in an arrest or summons?
:::


```{r}
#| eval: false
sqf_2011 |>
  #Subset to rows where suspect innocent
  _____(arrestsumm _____ 0) |> 
  #Calculate number of observations
  _____(total_innocent = n(), 
            percent_innocent = _____ / total_stops * 100)
```

###### Question

::: question

In how many stops were the individuals aged 14-24? In what percentage of stops were the individuals aged 14-24?
:::


```{r}
#| eval: false
sqf_2011 |>
  #Subset to rows where suspect age 14-24
  _____(age _____ 14 & age _____ 24) |> 
  #Calculate number of observations and percentage of observations
  _____(total_14_24 = _____, 
            percent_14_24 = n() / total_stops * 100)
```

###### Question

::: question

Fix the code below to calculate the currect number of stops for individuals 14-24. 

:::


```{r}
#| eval: false
total_stops_age_recorded <-
  sqf_2011 |>
  #Subset to rows where age is not 999
  _____(age _____ 999) |> 
  summarize(Count = n()) |>
  pull()

sqf_2011 |>
  filter(age >= 14 & age <= 24) |>
  summarize(total_14_24 = n(), 
            percent_14_24 = n() / total_stops_age_recorded * 100)
```

This still doesn't match the values we see on the website, but it does match the values we see in the NYCLU's 2011 report on Stop, Question, and Frisk data. This is typically when I would reach out to a representative at the NYCLU to inquire about the discrepancy.


###### Question

::: question

How many stops were there per race in 2011? What percentage of stops per race in 2011? Arrange by number of stops in descending order.
:::


```{r}
#| eval: false
total_stops_race_recorded <-
  sqf_2011 |>
  #Subset to rows where race_cat is not NA or "OTHER"
  filter(_____(race_cat) & race_cat _____ "OTHER") |> 
  summarize(Count = n()) |>
  pull()

sqf_2011 |>
  #Subset to rows where race_cat is not NA or "OTHER"
  _____(_____(race_cat) & race_cat _____ "OTHER") |> 
  #Group by race
  _____(race_cat) |> 
  #Calculate number of observations
  _____(stops = n(), 
            percent_stops = n() / total_stops_race_recorded * 100) |>
  #Sort by stops in descending order
  _____(_____(stops)) 
```

###### Question

::: question

In how many stops were the individuals identified as Latinx (i.e. "WHITE-HISPANIC" or "BLACK-HISPANIC")? In what percentage of stops were the individuals identified as Latinx?
:::


```{r}
#| eval: false
sqf_2011 |>
  #Subset to rows where race_cat is "WHITE-HISPANIC" or "BLACK-HISPANIC"
  _____(race_cat _____ c("WHITE-HISPANIC", "BLACK-HISPANIC")) |> 
  #Calculate number of observations
  _____(stops_Latinx = _____, 
            percent_Latinx = n() / total_stops_race_recorded * 100)
```

###### Question

::: question

What percentage of stops in 2011 resulted in a frisk per race? What percentage of stops in 2011 resulted in a weapon found per race? What percentage of stops in 2011 resulted in an arrest or summons per race? In your resulting data table, each row should be a race, and there should be columns for `stops`, `percent_stops`, `percent_frisked`, `percent_wpnfound` , and `percent_arrestsumm`.
:::


```{r}
#| eval: false
# Write code here. 
```

###### Question

::: question

Below, in 2-3 sentences, summarize what you learn from reviewing the summary statistics from the code above. What does this tell us about the constitutionality of 2011 stop and frisk activity in NYC?
:::



```{r}
#| document: show
#If you finish early, I encourage you to attempt to plot some of this data below using `ggplot()`!
```

