---
title: "Lab 5: Exploratory Data Analysis"
author: "YOUR NAME"
format: html
editor: 
  mode: source
bibliography: IntroDS.bib
---

## Setting Up Your Environment

```{r}
#| warning: false
#| messages: false
library(tidyverse)

sqf_url <- "https://www1.nyc.gov/assets/nypd/downloads/zip/analysis_and_planning/stop-question-frisk/sqf-2011-csv.zip"
temp <- tempfile()
download.file(sqf_url, temp)
sqf_zip <- unzip(temp, "2011.csv")
sqf_2011 <- read.csv(sqf_zip, stringsAsFactors = FALSE) 
sqf_2011_race_cat <- read.csv("https://raw.githubusercontent.com/lindsaypoirier/STS-101/master/Data/SQF/sqf_race_categories.csv", stringsAsFactors = FALSE) 
rm(sqf_url)
rm(temp)
rm(sqf_zip)
```

```{r}
sqf_2011 <- 
  sqf_2011 %>% 
  select(pct, race, age, frisked, pistol, riflshot, asltweap, knifcuti, machgun, othrweap, sumissue, arstmade) %>%
  left_join(sqf_2011_race_cat, by = "race") %>%
  mutate(across(frisked:arstmade, 
         ~ case_when(. == "Y" ~ 1, . == "N" ~ 0)))
rm(sqf_2011_race_cat)
```

::: callout-important
## Exercise 1

Add new columns indicating 1) whether a weapon was found or 2) an arrest/summons was made.
:::

```{r eval=FALSE}
sqf_2011 <- 
  sqf_2011 %>%
  #Add a variable for weapon found
  _____(wpnfound = case_when(pistol == 1 |
                               riflshot == 1 | 
                               asltweap == 1 |
                               knifcuti == 1 | 
                               machgun == 1 | 
                               othrweap == 1 ~ 1,
                             TRUE ~ 0))
sqf_2011 <- 
  sqf_2011 %>%
  #Add a variable for arrest made or summons issued
  _____(arrestsumm = case_when(sumissue == 1 | 
                                arstmade == 1 ~ 1,
                               TRUE ~ 0))
```

::: callout-important
## Exercise 2

Subset the dataset to the six variables listed in the data dictionary above.
:::

```{r eval=FALSE}
sqf_2011 <-
  sqf_2011 %>%
  _____(pct, arrestsumm, _____, wpnfound, race_cat, _____)
```

::: callout-important
## Exercise 3

Calculate the number of stops in 2011. Hint: Keep in mind that every row in the dataset represents one stop.
:::

```{r eval = FALSE}
total_stops <-
  sqf_2011 %>%
  summarize(Count = _____) %>%
  pull()

total_stops
```

::: callout-important
## Exercise 4

How many stops did not result in an arrest or summons in 2011? What percentage of stops did not result in an arrest or summons?
:::

```{r eval=FALSE}
sqf_2011 %>%
  #Subset to rows where suspect innocent
  _____(arrestsumm _____ 0) %>% 
  #Calculate number of observations
  _____(total_innocent = n(), 
            percent_innocent = _____ / total_stops * 100)
```

::: callout-important
## Exercise 5

In how many stops were the individuals aged 14-24? In what percentage of stops were the individuals aged 14-24?
:::

```{r eval=FALSE}
sqf_2011 %>%
  #Subset to rows where suspect age 14-24
  _____(age _____ 14 & age _____ 24) %>% 
  #Calculate number of observations and percentage of observations
  _____(total_14_24 = _____, 
            percent_14_24 = n() / total_stops * 100)
```

```{r eval=FALSE}
total_stops_age_recorded <-
  sqf_2011 %>%
  #Subset to rows where age is not 999
  _____(age _____ 999) %>% 
  summarize(Count = n()) %>%
  pull()

sqf_2011_sub %>%
  filter(age >= 14 & age <= 24) %>%
  summarize(total_14_24 = n(), 
            percent_14_24 = n() / total_stops_age_recorded * 100)
```

::: callout-important
## Exercise 6

How many stops were there per race in 2011? What percentage of stops per race in 2011? Arrange by number of stops in descending order.
:::

```{r eval=FALSE}
total_stops_race_recorded <-
  sqf_2011 %>%
  #Subset to rows where race_cat is not NA or "OTHER"
  filter(_____(race_cat) & race_cat _____ "OTHER") %>% 
  summarize(Count = n()) %>%
  pull()

sqf_2011 %>%
  #Subset to rows where race_cat is not NA or "OTHER"
  _____(_____(race_cat) & race_cat _____ "OTHER") %>% 
  #Group by race
  _____(race_cat) %>% 
  #Calculate number of observations
  _____(stops = n(), 
            percent_stops = n() / total_stops_race_recorded * 100) %>%
  #Sort by stops in descending order
  _____(_____(stops)) 
```

```{r eval=FALSE}
sqf_2011 %>%
  #Subset to rows where race_cat is "WHITE-HISPANIC" or "BLACK-HISPANIC"
  _____(race_cat _____ c("WHITE-HISPANIC", "BLACK-HISPANIC")) %>% 
  #Calculate number of observations
  _____(stops_Latinx = _____, 
            percent_Latinx = n() / total_stops_race_recorded * 100)
```

::: callout-important
## Exercise 7

What percentage of stops in 2011 resulted in a frisk per race? What percentage of stops in 2011 resulted in a weapon found per race? What percentage of stops in 2011 resulted in an arrest or summons per race? In your resulting data table, each row should be a race, and there should be columns for `stops`, `percent_stops`, `percent_frisked`, `percent_wpnfound` , and `percent_arrestsum`.
:::

```{r eval=FALSE}
# Write code here. 
```


